###used a function so it would take less space!!!!!
## Data sets:
```{r}
library(stargazer)
library(car)
require(graphics)
ND <- read.csv("New_DDAY.csv")
API <- read.csv("API_A.csv")
Pop_Growth <- read.csv("Pop_Growth.csv")
unemployment <- read.csv("Unemployment.csv")
Education <- read.csv("Education.csv")
```

## Cleaning the Data:
For all the data sets being presented here, all were downloaded in a wide form for an excel spread sheet. Any other method did not work, which is why for the idea to work I needed to have it in long form. So after stripping each row, I would divide each row into groups of 270 of 65 indexes long. Then i would be able to create a new data frame for each csv file so it be easier to ru further code and not look like an eyesore otherwise. The associated varible names were df_ND,df_API,df_PG,df_Unemp,df_ed.

# Code for ND

```{r}
swappy_swappy <- function(df){
  row <- c() 
  mid <- c()
  not_imp <- c() #useless vector
  for (ind_row in rownames(df)){
    for (i in df[ind_row,]){
      RC <- as.character(i)
      if (is.na(RC)){
        RC <- 0
        row <- c(row, RC)
        
       } else {
         vari_RC <- suppressWarnings(as.numeric(RC))
                                        
                                        
          if (!is.na(vari_RC)){
            row <- c(row, RC)
            }  else if (RC != ""){
            not_imp <- c(not_imp, RC)
            }
      }
    }
  }
  df_by_65 <- rep(1:270, each = 65)
  fin <- split(row, df_by_65)
  return(fin)
}

```

# Data frame cleaning for Poverty Rate:
```{r}
fin_ND <- swappy_swappy(ND)
col_nm_all <- ND[,1][5:270]
df_ND <- data.frame(fin_ND[5:270])
colnames(df_ND) <- c(col_nm_all)
years <- c(1960:2024) 
rownames(df_ND) <- c(years)
df_ND$idea <- ifelse(df_ND$Aruba >= 0, 0)
for (idk in colnames(df_ND)){
  df_ND$idea <- as.numeric(df_ND$idea) + as.numeric(df_ND[[idk]])
  df_ND$Poverty_rate <- ifelse(df_ND$Aruba >= 0, df_ND$idea/ 270)
}
```

# Data Frame cleaning for Inflation:
```{r}
fin_API <- swappy_swappy(API)

col_nm_2 <- API[,1][5:270]
df_API <- data.frame(col_nm = fin_API[5:270])

colnames(df_API) <- c(col_nm_2)
rownames(df_API) <- c(years)
df_API$idea <- ifelse(df_API$Aruba >= -1, 0,0)
for (idk_2 in colnames(df_API)){
  df_API$idea <- as.numeric(df_API$idea) + as.numeric(df_API[[idk_2]])
  df_ND$`Inflation` <- ifelse(df_API$Aruba >= -1, df_API$idea/ 270, 0)
}
```

#Data Farme Cleaning for Population growth:
```{r}
fin_PG <- swappy_swappy(Pop_Growth)

col_nm_3 <- Pop_Growth[,1][5:270]
df_PG <- data.frame(col_nm = fin_PG[5:270])

colnames(df_PG) <- c(col_nm_3)
rownames(df_PG) <- c(years)
df_PG$idea <- ifelse(df_PG$Aruba >= -1, 0,0)
for (idk_3 in colnames(df_API)){
  df_PG$idea <- as.numeric(df_PG$idea) + as.numeric(df_PG[[idk_3]])
  df_ND$`Population` <- ifelse(df_PG$Aruba >= -10000,0, df_PG$idea/ 270)
}
```

#Data Frame Cleaning for Unemployment
```{r}
fin_unempo <- swappy_swappy(unemployment)

col_nm_4 <- unemployment[,1][5:270]
df_unemployment <- data.frame(fin_unempo[5:270])

colnames(df_unemployment) <- c(col_nm_4)
rownames(df_unemployment) <- c(years)

df_unemployment$idea <- ifelse(df_unemployment$Aruba >= -1, 0,0)
for (idk_4 in colnames(df_unemployment)){
  df_unemployment$idea <- as.numeric(df_unemployment$idea) + as.numeric(df_unemployment[[idk_4]])
  df_ND$`Unemployment` <- ifelse(df_unemployment$Aruba >= -10000,df_unemployment$idea/ 270,0)
}
```

#Data Frame Cleaning for Education % from GDP
```{r}
fin_ed <- swappy_swappy(Education)

col_nm_ed <- Education[,1][5:270]
df_ed <- data.frame(col_nm_ed = fin_ed[5:270])

colnames(df_ed) <- c(col_nm_ed)
rownames(df_ed) <- c(years)
df_ed$Education <- ifelse(df_ed$Aruba >= -1, 0,0)
for (idk_ed in colnames(df_ed)){
  df_ed$Education <- as.numeric(df_ed$Education) + as.numeric(df_ed[[idk_ed]])
  df_ND$`Education % from GDP` <- ifelse(df_ed$Aruba >= -10000,df_ed$Education/ 270)
}
```
