##Question: Is there a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

---
title: 'STAT 107 Project:PHC'
author: "Marianne, Donnie, Uriel"
date: '2025-10-14'
output:
  pdf_document: default
  html_document: default
---
##Introduction (work in progress)
#Purpose:
To created and train a model to not only establish a connection between the Poverty headcount ratio and Inflation. But potential create a means to see trends that will eventually help foresee potential spikes of inflation worldwide. (work in process)

#Question
The question we are attempting to answer is whether or not there is a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

#Benefit:
This analysis should be able to benefit government institutions such Federal Reserves and IRS (Internal revenue Service) of the US but many more countries. (Add more)

#Plan:
We hope to answer our hypothesis since, if we are able to establish a connection between inflation to PHC. Then going forward it would be possible to train a model that could potential forecast the trends of inflation which would help government bodies worldwide create plans to counteract these shifts. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Main data set idea
 
```{r}

DDAY_1 <- read.csv("DDAY.csv")
head(DDAY_1)
summary(DDAY_1)
table(DDAY_1)


```
##Supporting data sets
```{r}
API_1 <- read.csv("API_A.csv")
head(API_1)

API_2 <- read.csv("API_B.csv")
head(API_2)
```
```{r}
Grow_1 <- read.csv("Grow.csv")
head(Grow_1)
summary(Grow_1)
```



## Basic Data Frame Descriptions


First questions!

1. How many observations: `nrow()` function.
2. How many variables: `ncol()` function.
3. Dimensions of data frame: `dim()` function.

```{r}

nrow(DDAY_1) # No. of observations.
ncol(DDAY_1) # No. of variables of interest.
dim(DDAY_1)  # both.

```


4. Names attached to each variable: `colnames()` function.

```{r}

colnames(DDAY_1)

```


5. Looking at the first 5 rows: `head()` function.

```{r}

head(DDAY_1)

```

