##Question: Is there a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

---
title: 'STAT 107 Project:PHC'
author: "Marianne, Donnie, Uriel"
date: '2025-10-14'
output:
  pdf_document: default
  html_document: default
---
##Introduction (work in progress)
#Purpose:
To created and train a model to not only establish a connection between the Poverty headcount ratio and Inflation. But potential create a means to see trends that will eventually help foresee potential spikes of inflation worldwide. (work in process)

#Question
The question we are attempting to answer is whether or not there is a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

#Benefit:
This analysis should be able to benefit government institutions such Federal Reserves and IRS (Internal revenue Service) of the US but many more countries. (Add more)

#Plan:
We hope to answer our hypothesis since, if we are able to establish a connection between inflation to PHC. Then going forward it would be possible to train a model that could potential forecast the trends of inflation which would help government bodies worldwide create plans to counteract these shifts. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Main data set idea
The data that is available is DDAY (Poverty Head Count at $3.00 a day (2021 PPP)), API_1 and API_2 are from the same data set (Inflation consumer prices (annual %) for the World), Grow (Population growth (annual%)).
 
```{r}

ND <- read.csv("New_DDAY.csv")
## Since none of the actual function for r would work with any of the data sets since all of them end up using the column values.
#swap columns -> rows by making a new data frame
```

```{r}
col_ND <- ND[,1][5:270]
#code above gets all the column names into a vector

row_ND <- c()
not_imp <- c()
for (ind_row in 1:nrow(ND)){
  ND[is.na(ND)] <- 0
  for (i in ND[ind_row,]){
    RC <- as.character(i)
    if (RC == 0){
      row_ND <- c(row_ND, RC)
        
      } else {
        vari_RC <- suppressWarnings(as.numeric(RC))
                                    
                                    
      if (!is.na(vari_RC)){
        row_ND <- c(row_ND, RC)
        }  else if (RC != ""){
        not_imp <- c(not_imp, RC)
        }
    }
  }
}
```

###Explanation for why I'm doing this
since I used the loops to get all the rows into a singular vector, and after running length(row_ND) I know it is 17550 spaces long. So there are 270 rows for every data set, and we want them to be separate by 65 each. Because each space will have an associated batch of numbers so it would match up right with the years (1960-2025). 
Please use this explanation for the remainder two data sets.
```{r}
x <- row_ND
ND_by_65 <- rep(1:270, each = 65)
fin_ND <- split(x, ND_by_65)
fin_ND[4]
fin_ND[270]

```


#Making the data frame
```{r}

```




##Supporting data sets
```{r}
API <- read.csv("API_A.csv")
SN_API <- read.csv("API_B.csv")
#This just has notes left behind by the orginal makers of this dataset
```


```{r}
row_API <- c()
not_imp_2 <- c()
for (ind_API in 1:nrow(API)){
  API[is.na(API)] <- 0
  for (i in API[ind_API,]){
    RC_2 <- as.character(i)
    if (RC_2 == 0){
      row_API <- c(row_API, RC_2)
        
        } else {
        vari_RC_2 <- suppressWarnings(as.numeric(RC_2))
                                    
                                    
        if (!is.na(vari_RC_2)){
          row_API <- c(row_API, RC_2)
          } else if (RC_2 != ""){
          not_imp_2 <- c(not_imp_2, RC_2)
        }
    }
  }
}
```


```{r}
x <- row_API
API_by_65 <- rep(1:270, each = 65)
fin_API <- split(x, API_by_65)
fin_API[4]
fin_API[270]
#works
```


```{r}
Pop_Growth <- read.csv("Pop_Growth.csv")
```

```{r}
row_PG <- c()
NI <- c()
for (ind_PG in 1:nrow(Pop_Growth)){
  Pop_Growth[is.na(Pop_Growth)] <- 0
  for (i in Pop_Growth[ind_PG,]){
    RC_3 <- as.character(i)
    if (RC_3 == "0"){
      row_PG <- c(row_PG, RC_3)
        
      } else {
      vari_RC_3 <- suppressWarnings(as.numeric(RC_3))
                                    
                                    
      if (!is.na(vari_RC_3)){
        row_PG <- c(row_PG, RC_3)
      } else if (RC_3 != ""){
        NI <- c(NI, RC_3)
        }
    }
  }
}
```

```{r}
x <- row_PG
PG_by_65 <- rep(1:270, each = 65)
fin_PG <- split(x, PG_by_65)
fin_PG[4]
fin_PG[270]
#works
```


## Basic Data Frame Descriptions


First questions!

1. How many observations: `nrow()` function.
2. How many variables: `ncol()` function.
3. Dimensions of data frame: `dim()` function.

```{r}

nrow(ND) # No. of observations.
ncol(ND) # No. of variables of interest.
dim(ND)  # both.

```


4. Names attached to each variable: `colnames()` function.

```{r}
# if we want column names, we have to find their columns
ND[4,5:66]
ND[2,1]

```


5. Looking at the first 5 rows: `head()` function.

```{r}

head(ND)

```

