##Question: Is there a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

---
title: 'STAT 107 Project:PHC'
author: "Marianne, Donnie, Uriel"
date: '2025-10-14'
output:
  pdf_document: default
  html_document: default
---
##Introduction (work in progress)
#Purpose:
To created and train a model to not only establish a connection between the Poverty headcount ratio and Inflation. But potential create a means to see trends that will eventually help foresee potential spikes of inflation worldwide. (work in process)

#Question
The question we are attempting to answer is whether or not there is a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

#Benefit:
This analysis should be able to benefit government institutions such Federal Reserves and IRS (Internal revenue Service) of the US but many more countries. (Add more)

#Plan:
We hope to answer our hypothesis since, if we are able to establish a connection between inflation to PHC. Then going forward it would be possible to train a model that could potential forecast the trends of inflation which would help government bodies worldwide create plans to counteract these shifts. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Main data set idea
The data that is available is DDAY (Poverty Head Count at $3.00 a day (2021 PPP)), API_1 and API_2 are from the same data set (Inflation consumer prices (annual %) for the World), Grow (Population growth (annual%)).
 
```{r}

ND <- read.csv("New_DDAY.csv")
head(ND, n =10)
summary(ND)


```
##Supporting data sets
```{r}
API <- read.csv("API_A.csv")
print(API[5,])
#using the summary() function gives a output, but since the table itself uses rows as columns and vice-versa. The outputs are mathematically incorrect
summary(API[5,])
# This SN_API is just notes left behind by the people who actually made the data set, they might prove to be useful later on when it comes to explaining the data.
SN_API <- read.csv("API_B.csv")
head(SN_API)
```
```{r}
Pop_Growth <- read.csv("Pop_Growth.csv")
head(Pop_Growth)
#look at previous chunk for explanation
summary(Pop_Growth)
```



## Basic Data Frame Descriptions


First questions!

1. How many observations: `nrow()` function.
2. How many variables: `ncol()` function.
3. Dimensions of data frame: `dim()` function.

```{r}

nrow(ND) # No. of observations.
ncol(ND) # No. of variables of interest.
dim(ND)  # both.

```


4. Names attached to each variable: `colnames()` function.

```{r}
# if we want column names, we have to find their columns
ND[1]

```


5. Looking at the first 5 rows: `head()` function.

```{r}

head(ND)

```

