##Question: Is there a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

---
title: 'STAT 107 Project:PHC'
author: "Marianne, Donnie, Uriel"
date: '2025-10-14'
output:
  pdf_document: default
  html_document: default
---
##Introduction (work in progress)
#Purpose:
To created and train a model to not only establish a connection between the Poverty headcount ratio and Inflation. But potential create a means to see trends that will eventually help foresee potential spikes of inflation worldwide. (work in process)

#Question
The question we are attempting to answer is whether or not there is a statistically significant correlation between inflation rates and the Poverty headcount ratio across countries worldwide?

#Benefit:
This analysis should be able to benefit government institutions such Federal Reserves and IRS (Internal revenue Service) of the US but many more countries. (Add more)

#Plan:
We hope to answer our hypothesis since, if we are able to establish a connection between inflation to PHC. Then going forward it would be possible to train a model that could potential forecast the trends of inflation which would help government bodies worldwide create plans to counteract these shifts. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Main data set idea
The data that is available is DDAY (Poverty Head Count at $3.00 a day (2021 PPP)), API_1 and API_2 are from the same data set (Inflation consumer prices (annual %) for the World), Grow (Population growth (annual%)).
 
```{r}
ND <- read.csv("New_DDAY.csv")
```

```{r}
#Item names: row_ND, not_imp, ind_row, RC, row_ND,vari_RC.
row_ND <- c() 
not_imp <- c() #useless vector
for (ind_row in rownames(ND)){
  for (i in ND[ind_row,]){
    RC <- as.character(i)
    if (is.na(RC)){
      RC <- 0
      row_ND <- c(row_ND, RC)
      
     } else {
       vari_RC <- suppressWarnings(as.numeric(RC))
                                      
                                      
        if (!is.na(vari_RC)){
          row_ND <- c(row_ND, RC)
          }  else if (RC != ""){
          not_imp <- c(not_imp, RC)
          }
    }
  }
}
```

# Get groups of 270 and each group has a length of 65. Then use split() to apply that row_ND
```{r}
a <- row_ND
ND_by_65 <- rep(1:270, each = 65)
fin_ND <- split(a, ND_by_65)

```


## Data Frame for ND with rows and columns switched: df_ND
```{r}
# must have this chunk running first if you want the other similar section to function!!
col_nm <- ND[,1][5:270]
df_ND <- data.frame(fin_ND[5:270])

colnames(df_ND) <- c(col_nm)
years <- c(1960:2024) 
rownames(df_ND) <- c(years)
```

##Basic R functions:
```{r}
##df_ND[,266] is the last country (Zimbabwe). Because I set fin_ND[5:270], so we can skip the first 3 rows of zeros and country years
i = 1
while ( i < 267 ){
  #print(i)
  #print(summary(as.double(df_ND[,i])))
  #have a pound sign so not to clogged the file
  i = i + 1
  
}
#Used a while loop since it was better option.

```





##Supporting data sets
```{r}
API <- read.csv("API_A.csv")
SN_API <- read.csv("API_B.csv")
#This just has notes left behind by the original makers of this csv file.
```


```{r}
#Item names: row_API, not_imp_2, ind_API, RC_2,vari_RC_2.
row_API <- c() 
not_imp_2 <- c() #useless vector
for (ind_API in rownames(API)){
  for (i in API[ind_API,]){
    RC_2 <- as.character(i)
    if (is.na(RC_2)){
      RC_2 <- 0
      row_API <- c(row_API, RC_2)
      
     } else {
       vari_RC_2 <- suppressWarnings(as.numeric(RC_2))
                                    
                                    
        if (!is.na(vari_RC_2)){
          row_API <- c(row_API, RC_2)
          } else if (RC_2 != ""){
          not_imp_2 <- c(not_imp_2, RC_2)
        }
    }
  }
}
```


```{r}
b <- row_API
API_by_65 <- rep(1:270, each = 65)
fin_API <- split(b, API_by_65)
```

##Data Frame for API with rows and columns switched: df_API
```{r}
col_nm_2 <- API[,1][5:270]
df_API <- data.frame(col_nm = fin_API[5:270])

colnames(df_API) <- c(col_nm_2)
rownames(df_API) <- c(years)

i = 1
while ( i < 267 ){
  #(i)
  #print((summary(as.double(df_API[,i]))))
  #have a pound sign so not to clogged the file
  i = i + 1
  
}
```


```{r}
Pop_Growth <- read.csv("Pop_Growth.csv")
```

```{r}
#Item names: row_PG, NI, ind_PG, RC_3,vari_RC_3.
row_PG <- c()
NI <- c() #useless vector
for (ind_PG in rownames(Pop_Growth)){
  for (i in Pop_Growth[ind_PG,]){
    RC_3 <- as.character(i)
    if (is.na(RC_3)){
      RC_3 <- 0
      row_PG <- c(row_PG, RC_3)
      
     } else {
       vari_RC_3 <- suppressWarnings(as.numeric(RC_3))
                                    
                                    
                                    
      if (!is.na(vari_RC_3)){
        row_PG <- c(row_PG, RC_3)
      } else if (RC_3 != ""){
        NI <- c(NI, RC_3)
        }
    }
  }
}
```

#Look at main data set grouping section for explanation
```{r}
c <- row_PG
PG_by_65 <- rep(1:270, each = 65)
fin_PG <- split(c, PG_by_65)
```

##Data Frame for Pop_Growth with rows and columns switched: df_PG
```{r}
col_nm_3 <- Pop_Growth[,1][5:270]
df_PG <- data.frame(col_nm = fin_PG[5:270])

colnames(df_PG) <- c(col_nm_3)
rownames(df_PG) <- c(years)

i = 1
while ( i < 267 ){
  #(i)
  #print((summary(as.double(df_PG[,i]))))
  #have a pound sign so not to clogged the file
  i = i + 1
  
}
```


## Basic Data Frame Descriptions


First questions!

1. How many observations: `nrow()` function.
2. How many variables: `ncol()` function.
3. Dimensions of data frame: `dim()` function.

```{r}

nrow(df_ND) # No. of observations.
ncol(df_ND) # No. of variables of interest.
dim(df_ND)  # both.

```


4. Names attached to each variable: `colnames()` function.

```{r}
# if we want column names, we have to find their columns
ND_col <- colnames(df_ND)
ND_row <- rownames(df_ND)

```


5. Looking at the first 5 rows: `head()` function.

```{r}

head(df_ND)

```

